---
title: "vi"
author: "Adit Doshi, Neha Vanjara, Yash Kharade"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,fig.align="center",echo = FALSE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```


```{r Loading the required packages}
packages <-
  c(
    "tidyverse",
    "janitor",
    "data.table",
    "devtools",
    "ggmap",
    "viridis",
    "lubridate",
    "reshape",
    "ggplot2",
    "wordcloud",
    "knitr",
    "rworldmap",
    "rworldxtra",
    "dplyr",
    "tidyr",
    "kableExtra"
  )
install.packages(setdiff(packages, rownames(installed.packages())))
for (package in packages) {
  library(package, character.only = TRUE)
}
```




```{r Loading_data}
# Loading NYC311 DataSet
#df <- fread("311_Service_Requests_from_2010_to_Present.csv")
nyc311 <- fread("311_Service_Requests_from_2010_to_Present.csv") 
names(nyc311)<-names(nyc311) %>% stringr::str_replace_all("\\s", "")# Removing the space

```

```{r}
nyc311<-select(nyc311,-c(FerryTerminalName,FerryDirection,GarageLotName,BridgeHighwaySegment,
                 RoadRamp,BridgeHighwayDirection,BridgeHighwayName,TaxiPickUpLocation,
                 TaxiCompanyBorough,VehicleType,SchoolorCitywideComplaint,SchoolNotFound,
                 SchoolZip,SchoolState,SchoolState,SchoolCity,SchoolAddress,SchoolPhoneNumber,
                 SchoolCode,SchoolRegion,SchoolNumber,SchoolName,ParkBorough,ParkFacilityName
                 ,CommunityBoard,ResolutionActionUpdatedDate,IntersectionStreet2,IntersectionStreet1,
                 CrossStreet2,CrossStreet1,))
```

```{r}

abc <- dplyr:: select(nyc311, Agency)

abc$Frequency <- ave(seq_len(nrow(abc)),abc,FUN=length)

abc <- abc[!duplicated(abc),]

abc <- abc[order(-abc$Frequency)]

abc <- abc %>% slice(1:11)



abc$Count <- abc$Frequency/1000

ggplot(data=abc)+geom_bar(mapping=aes(x=Agency,y=Count,fill=Count),stat = "identity")+
  ggtitle("Agencies handeling cases(frequency is in thousands): ")+
  xlab("Agencies")+ylab(" Complaints Frequency")+
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(color="black", size=20, face="bold"),
axis.title.x = element_text(color="black", size=10),
axis.title.y = element_text(color="black", size=10))
  

```



```{r top30}
#1 Top 30 Complaint Types
ggplot(data=subset(nyc311,ComplaintType %in% count(nyc311,ComplaintType,
                                               sort = TRUE)
                   [1:30,]$ComplaintType),
       aes(x=ComplaintType))+
  geom_bar(fill='darkgreen')+
  labs(x='Complaint Type',y='Number of Complaints')+
  coord_flip()+ggtitle("Plot of Complaint Type VS Number Of Complaints")+theme(
    plot.title = element_text(color="black", size=14, face="bold"),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14)) + 
  theme(plot.title=element_text(hjust=0.5))
```



```{r}
#2 Map Graph
dataset_map <- subset(nyc311,ComplaintType %in% count(nyc311,ComplaintType,sort = TRUE)[1:5,]$ComplaintType)
library(plyr)
dataset_map <- dataset_map %>% select(ComplaintType, Borough, Latitude, Longitude) %>% drop_na()
counts <- ddply(dataset_map, .(ComplaintType), "count")
filtered <- filter(counts, freq > 150)
filtered$freq <- as.numeric(filtered$freq)
filtered$Longitude <- as.numeric(filtered$Longitude)
filtered$Latitude <- as.numeric(filtered$Latitude)
FirstMap <- getMap(resolution = "high")
nyc_map <- ggplot() + geom_polygon(data = FirstMap, aes(x=long, y = lat, group = group),
                                   fill = "white", color = "gray") + xlim(-74.5, -73.5)+
  ylim(40.5, 41)
dataset_map <- dataset_map %>% select(ComplaintType, Borough, Latitude, Longitude) %>% drop_na()
counts <- ddply(dataset_map, .(ComplaintType), "count")
filtered <- filter(counts, freq > 150)
filtered$freq <- as.numeric(filtered$freq)
filtered$Longitude <- as.numeric(filtered$Longitude)
filtered$Latitude <- as.numeric(filtered$Latitude)
FirstMap <- getMap(resolution = "high")
nyc_map <- ggplot() + geom_polygon(data = FirstMap, aes(x=long, y = lat, group = group),
                                   fill = "white", color = "gray") + xlim(-74.5, -73.5)+
  ylim(40.5, 41) 

## Regions defined for each Polygons
nyc_map + geom_point(data=filtered, aes(Longitude, Latitude, size=freq), colour="red") +
  facet_wrap(~ComplaintType, scales = "free") +
  labs(x = "Longitude", y = "Latitude", title = "Density of complaints by complaint type") +
  scale_size(name="Number of Complaints")+ theme(plot.title = element_text(size=25)) +
  theme(plot.title=element_text(hjust=0.5))
detach("package:plyr")
```



```{r}
#3 Complaint Types Borough Wise
heatmap<- select(nyc311, Borough,ComplaintType)
xtabA<- filter(heatmap,
                     ComplaintType %in%c('Water System','Traffic Signal Condition',
                                         'Street Light Condition','Street Condition',
                                         'PLUMBING','PAINT - PLASTER','NONCONST',
                                         'HEATING',
                                         'HEAT/HOT WATER','GENERAL CONSTRUCTION',
                                         'Blocked Driveway'))

xtabA<- filter(xtabA, Borough %in%c('BRONX','BROOKLYN','MANHATTAN','QUEENS','STATEN ISLAND'))
p<-table(xtabA$Borough,xtabA$ComplaintType)
data_melt<-melt(p)
xc<-ggplot(data_melt,aes(x=Var.1,y=Var.2))+geom_tile(aes(fill=value))+
  scale_fill_gradient(low="yellow",high="red")
xc+ggtitle("Heatmap for top 11 complaints: ")+xlab("Borough")+ylab("Complaint Type")+theme(
  plot.title = element_text(color="black", size=14, face="bold"),
  axis.title.x = element_text(color="black", size=14),
  axis.title.y = element_text(color="black", size=14)
) + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + theme(plot.title=element_text(hjust=0.5))
```


```{r}
#Heating Complaint type:
calls <- nyc311 %>% filter(ComplaintType == "HEATING") %>%
  mutate(standard_date = as.Date(substring(CreatedDate, 1, 10), '%m/%d/%Y')) %>%
  mutate(month_of_year = months(standard_date))
months_table = table(calls$month_of_year)
months_table = months_table[c(5, 4, 8, 1, 9, 7, 6, 2, 12, 11, 10, 3)]
months_table <- data.frame(months_table)
colnames(months_table) <- c("month_of_year", "No_of_calls")

table<-months_table
colnames(table) <- c("Month", "Heating calls")

table <-table %>%  kbl(caption = "Month wise heating calls:") %>% 
  kable_classic(full_width = F, html_font = "Cambria")

table
```



```{r}
b<- ggplot(months_table, aes(x=month_of_year)) +
  geom_line( mapping = aes(y=No_of_calls,group =1),color="blue") + 
  ggtitle("Number of calls month wise")+theme(
    plot.title = element_text(color="black", size=14, face="bold")) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(plot.title=element_text(hjust=0.5))
b
```




```{r}
Heating <- nyc311 %>% filter(ComplaintType %in%c("HEATING")) %>%
  mutate(standard_date = as.Date(substring(CreatedDate, 1, 10), '%m/%d/%Y')) 
  

Heating$Month_Yr <- format(as.Date(Heating$standard_date), "%Y-%m")

Heating <- Heating %>% group_by(Month_Yr)
Heating <- dplyr:: select(Heating, Month_Yr)
Heating <- as.data.frame(table(Heating))


ggplot(data=Heating)+geom_bar(mapping=aes(x=Month_Yr,y=Freq,fill=Freq), stat="identity")+
  ggtitle("Heating complaints by month: ")+xlab("Month")+ylab("Heating Complaints Frequency")+
  theme(axis.text.x = element_text(angle = 90),
axis.title.x = element_text(color="black", size=10),
axis.title.y = element_text(color="black", size=10))


```



```{r}
df2<-fread("https://data.ny.gov/api/views/krt9-ym2k/rows.csv?accessType=DOWNLOAD&sorting=true")

library(purrr)
names(df2)
colnames(df2)[2] <- "Borough"
df2$Borough <-replace(df2$Borough, df2$Borough=="New York County", "MANHATTAN")
df2$Borough <- replace(df2$Borough, df2$Borough=="Bronx County", "BRONX")
df2$Borough <- replace(df2$Borough, df2$Borough=="Kings County", "BROOKLYN")
df2$Borough <- replace(df2$Borough, df2$Borough=="Richmond County", "STATEN ISLAND")
df2$Borough <- replace(df2$Borough, df2$Borough=="Queens County", "QUEENS")

names(df2) <- names(df2) %>%
  stringr::str_replace_all("\\s", "")

df2<-dplyr::filter(df2,ProgramType %in%
                     c('Postcensal Population Estimate','Intercensal Population Estimate'))

df2<-dplyr::select(df2,Borough,Year,Population)
colnames(df2)[1] <- "Borough"

df2 <- df2 %>% 
  select(Borough, Year, Population) %>%  
  filter(Borough %in% c("BROOKLYN", "BRONX", "MANHATTAN", "STATEN ISLAND", "QUEENS")) %>%
  filter(Year %in% c("2005","2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014"))

df2$Year <- as.character(df2$Year)

capita <- select(nyc311,c(ComplaintType,CreatedDate,Borough))
capita <- capita %>% filter(ComplaintType %in%c("HEATING"))
capita <- capita %>%mutate(CreatedDate = mdy_hms(CreatedDate))
capita$Year <- format(capita$CreatedDate, format = "%Y")
capita<-subset(capita,select=c(Borough,Year))
capita <- capita %>% filter(Year %in%c("2010","2011","2012","2013"))
capita <- capita %>% group_by(Year ,Borough)%>%mutate(count=n())
capita <- capita %>% distinct()


Heating_capita <- merge(df2, capita, by.x=c("Year", "Borough"), by.y=c("Year", "Borough"))
Heating_capita$per_capita <- Heating_capita$count/Heating_capita$Population

Bronx <- Heating_capita %>% filter(Borough %in%c("BRONX"))

Brooklyn <- Heating_capita %>% filter(Borough %in%c("BROOKLYN"))

Manhattan <- Heating_capita %>% filter(Borough %in%c("MANHATTAN"))

Queens <- Heating_capita %>% filter(Borough %in%c("QUEENS"))

Staten_Island <- Heating_capita %>% filter(Borough %in%c("STATEN ISLAND"))

plot(Bronx$Year,Bronx$per_capita, type='l',col='red',xlab="Year",ylab="Calls",
     main="Per Capita Heating Calls")
lines(Brooklyn$Year,Brooklyn$per_capita,col='blue')
lines(Manhattan$Year, Manhattan$per_capita,col='purple')
lines(Queens$Year, Queens$per_capita,col='green')
lines(Staten_Island$Year, Staten_Island$per_capita,col='black')
legend("topright",legend=c("Bronx","Brooklyn","Manhattan","Queens","Staten_Island"),
       col = c('red','blue','purple','green','black'),lty = 1)
```

This Figure shows 311 per capita calls for Heating related complaints. It is intresting to see that while all the borough's per capita calls are increasing, Bronx region is growing at a higher rate and Staten Island has the lowest growth rate. 
The visualization shows us the evolution of heating complaints per person in each borough over the years 2010 to 2013. The per capita calls for all boroughs increased at the maximum rate during the first year and then slowing down as the years pass.




```{r}
# cloud for Bronx heating complaint
cloud<-filter(nyc311,Borough %in%c('BRONX'))
cloud<-select(cloud,Borough,ComplaintType)
cloud <- cloud %>% group_by(ComplaintType)%>% summarise(Count = n()) %>% arrange(desc(Count))
wordcloud(words=cloud$ComplaintType,freq = cloud$Count,scale=c(4,.5),
          max.words=100,rot.per=.5,colors=palette())

```


```{r}
nyc311C <- nyc311 %>%
  mutate(ClosedDate = mdy_hms(ClosedDate)) %>%
  mutate(CreatedDate = mdy_hms(CreatedDate)) %>%
  mutate(DueDate = mdy_hms(DueDate))


nyc311C$Year <- format(nyc311C$CreatedDate, format = "%Y")


df1 <- nyc311C %>%
  select(Borough, Agency, ComplaintType, IncidentZip, CreatedDate, ClosedDate, Year,) %>%
  filter(Borough %in% c("BROOKLYN", "BRONX", "MANHATTAN", "STATEN ISLAND", "QUEENS")) %>%
  filter(Year %in% c("2005","2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015"))


newnyc311 <- inner_join(df1, df2, by = c("Borough", "Year"))
a<- newnyc311 %>% filter(Borough %in% c("BRONX")) %>% group_by(Year, Population) %>% summarize(Freq=n())

ggplot(data=a[1:10, ], mapping = aes(x=Year))+
  geom_line( mapping = aes(y=Population,group =1),color="blue")+
  geom_line( mapping = aes(y=Freq,group =2),color="black")+
  ggtitle("Trendline of Population & No. Of cases over the Years ")+xlab("Years")+ylab("Count")+
  theme(plot.title = element_text(color="black", size=14, face="bold"),
        axis.title.x = element_text(color="black", size=14),
        axis.title.y = element_text(color="black", size=14))
```


```{r}
sessionInfo()
```




